Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2016-01-26 cordillera-cycle.tex
Update sensitivity tests to 0.05 K simulations.

      As a~direct result of the different scaling factors applied
      (Table~\ref{tab:sens_params}), the resulting simulations show very
      little difference in the modelled glaciated areas corresponding to
      the maximum stage during MIS~2, but also during MIS~4
      (Table~\ref{tab:sens_extrema}). However, such can not be said of the
      modelled glaciated area corresponding to minimum stage during
      MIS~3. In fact, the extent of the remnant ice cap which persists over
      the Skeena Mountains during this stage shows a~significant sensitivity
      to ice rheology of 31\,\unit{\%}, and an even more important
      sensitivity to basal sliding parameters of 62\,\unit{\%}
      (Table~\ref{tab:sens_extrema}).

      The modelled sea-level potentials show more variability than
      the modelled glaciated areas (Table~\ref{tab:sens_extrema},
      Fig.~\ref{fig:sens_ts}). As one could expect, softer ice and weaker
      till both result in a~thinner ice sheet, while harder ice and stronger
      till result in a~thicker ice sheet. For instance, peak sea-level
      relevant ice volume
      during the MIS~2 (LGM) varies by 30\,\unit{\%} between the two
      parametrisations of ice rheology used, and by 21\,\unit{\%} between
      the two parametrisations of basal sliding used. The differences in
      ice volume are greatest during the MIS~3
      (Table~\ref{tab:sens_extrema}, Fig.~\ref{fig:sens_ts}) where both the
      areal and thickness contributions add up.

@@

      During MIS~2, the modelled sea level potential peaks at
      8.62\,\unit{m}\,s.l.e. (19.1\,\unit{ka}) in the GRIP simulation and at
      8.56\,\unit{m}\,s.l.e. (17.4\,\unit{ka}) in the EPICA simulation.
      However, these numers are subject to uncertainties on physics
      embedded in the model. The range of parameter values tested in our
      sensitivity study on the lower-resolution GRIP simulation
      (Sect.~\ref{sec:sens}) yielded relative errors of 30\,\unit{\%}
      regarding ice rheological parameters and 21\,\unit{\%} regarding basal
      sliding parameters (Fig.~\ref{fig:sens_ts};
      Table~\ref{tab:sens_extrema}).

@@

      During MIS~4, the modelled sea level potential peaks at
      7.43\,\unit{m}\,s.l.e. (57.6\,\unit{ka}) in the GRIP simulation and at
      4.84\,\unit{m}\,s.l.e. (61.9\,\unit{ka}) in the EPICA simulation,
      corresponding to respectively 86 and 57\,\unit{\%} of modelled MIS~2
      ice volumes.
      These estimates appear little sensitive to the range of parameter
      values tested in our sensitivity study (Sect.~\ref{sec:sens}), which
      yielded relative errors of 17\,\unit{\%} to ice rheological parameters
      and 21\,\unit{\%} to basal sliding parameters
      (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

      The modelled sea level potential corresponding to these persistent
      ice-dispersal centres attains a~minimum of 1.54\,\unit{m}\,s.l.e.
      (42.9\,\unit{ka}) in the GRIP simulation and of 2.55\,\unit{m}\,s.l.e.
      (52.4\,\unit{ka}) in the EPICA simulation, corresponding to
      respectively 18 and 30\,\unit{\%} of the MIS~2 ice volumes.
      However, these numbers should be considered with caution as our
      sensitivity study (Sect.~\ref{sec:sens}) shows that minimum ice volume
      during MIS~3 is highly sensitive to ice flow parameters, with relative
      errors of 57\,\unit{\%} to the range of ice rheological parameters
      tested and 120\,\unit{\%} to the range of basal sliding parameters
      tested (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

\begin{table*}
\caption{%
      Parameter values used in the sensitivity test.}
\label{tab:sens_params}
%\centering\makebox[\textwidth]
{\begin{tabular}{l|ccccc|cc|cc}
    \tophline
            & \multicolumn{5}{c|}{Rheology}
            & \multicolumn{2}{c|}{Sliding}
            & \multicolumn{2}{c}{GRIP scaling} \\
    Config. & $A_{\mathrm{c}}$ & $A_{\mathrm{w}}$
            & $Q_{\mathrm{c}}$ & $Q_{\mathrm{w}}$
            & $E_{\text{SIA}}$ & $\delta$ & $W_{\text{max}}$
            & $f$ & $T_{[32, 22]}$ \\
            & \multicolumn{2}{c}{(\unit{Pa^{-3}\,s^{-1}})}
            & \multicolumn{2}{c}{(\unit{J\,mol^{-1}})}
            & & & (\unit{m}) \\
    \middlehline
    Default$^1$  & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.02 & 2 & 0.38 & 6.2 \\
    \cline{1-10}
    Soft ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 5 & 0.02 & 2 & 0.40 & 6.6 \\
    Hard ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 1 & 0.02 & 2 & 0.37 & 6.0 \\
    \cline{1-10}
    Soft bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.01 & 1 & 0.40 & 6.5 \\
    Hard bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.05 & 5 & 0.36 & 5.9 \\
    \bottomhline
\end{tabular}}
\belowtable{%
      After $^1$\citet{Paterson.Budd.1982,Bueler.Pelt.2015};
      and $^2$\citet{Cuffey.Paterson.2010}.}
\end{table*}

@@

\begin{table*}[p]
\caption{%
      Extremes in Cordilleran ice sheet grounded ice extent and sea-level
      relevant ice volume corresponding to
      MIS~4, 3 and 2 using the GRIP paleo-climate forcing with each
      parameter configuration (Fig.~3). Relative differences (R. diff.) give
      rough error estimates related to varying selected ice rheology and
      basal sliding parameters (Table~\ref{tab:sens_params}).}
\label{tab:sens_extrema}
%\centering\makebox[\textwidth]
{\begin{tabular}{l*{3}{|ccc}}
    \tophline
             & \multicolumn{3}{c|}{Age (ka)}
             & \multicolumn{3}{c|}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m\,s.l.e.)} \\
    Config.  &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    Default  &  57.59 &  42.91 &  19.14
             &   1.93 &   0.67 &   2.09
             &   7.43 &   1.54 &   8.62 \\
    \cline{1-10}
    Soft ice &  58.89 &  49.97 &  21.57
             &   1.96 &   0.54 &   2.08
             &   6.58 &   1.03 &   6.88 \\
    Hard ice &  57.32 &  42.90 &  19.14
             &   1.90 &   0.75 &   2.12
             &   7.83 &   1.91 &   9.46 \\
    R. diff. &    3\,\unit{\%} &   16\,\unit{\%} &   13\,\unit{\%}
             &    3\,\unit{\%} &   31\,\unit{\%} &    2\,\unit{\%}
             &   17\,\unit{\%} &   57\,\unit{\%} &   30\,\unit{\%} \\
    \cline{1-10}
    Soft bed &  58.90 &  49.21 &  19.53
             &   1.88 &   0.55 &   2.05
             &   6.46 &   1.03 &   7.52 \\
    Hard bed &  57.31 &  42.91 &  19.14
             &   1.93 &   0.96 &   2.13
             &   7.99 &   2.89 &   9.31 \\
    R. diff. &    3\,\unit{\%} &   15\,\unit{\%} &    2\,\unit{\%}
             &    3\,\unit{\%} &   62\,\unit{\%} &    4\,\unit{\%}
             &   21\,\unit{\%} &  120\,\unit{\%} &   21\,\unit{\%} \\
    \bottomhline
\end{tabular}}
\end{table*}

>>>>

      As a~direct result of the different scaling factors applied
      (Table~\ref{tab:sens_params}), the resulting simulations show very
      little difference in the modelled glaciated areas corresponding to
      the maximum stage during MIS~2, but also during MIS~4
      (Table~\ref{tab:sens_extrema}). However, such can not be said of the
      modelled glaciated area corresponding to minimum stage during
      MIS~3. In fact, the extent of the remnant ice cap which persists over
      the Skeena Mountains during this stage shows a~significant sensitivity
      to ice rheology of 25\,\unit{\%}, and an even more important
      sensitivity to basal sliding parameters of 57\,\unit{\%}
      (Table~\ref{tab:sens_extrema}).

      The modelled sea-level potentials show more variability than
      the modelled glaciated areas (Table~\ref{tab:sens_extrema},
      Fig.~\ref{fig:sens_ts}). As one could expect, softer ice and weaker
      till both result in a~thinner ice sheet, while harder ice and stronger
      till result in a~thicker ice sheet. For instance, peak sea-level
      relevant ice volume
      during the MIS~2 (LGM) varies by 25\,\unit{\%} between the two
      parametrisations of ice rheology used, and by 14\,\unit{\%} between
      the two parametrisations of basal sliding used. The differences in
      ice volume are greatest during the MIS~3
      (Table~\ref{tab:sens_extrema}, Fig.~\ref{fig:sens_ts}) where both the
      areal and thickness contributions add up.

@@

      During MIS~2, the modelled sea level potential peaks at
      8.62\,\unit{m}\,s.l.e. (19.1\,\unit{ka}) in the GRIP simulation and at
      8.56\,\unit{m}\,s.l.e. (17.4\,\unit{ka}) in the EPICA simulation.
      However, these numers are subject to uncertainties on physics
      embedded in the model. The range of parameter values tested in our
      sensitivity study on the lower-resolution GRIP simulation
      (Sect.~\ref{sec:sens}) yielded relative errors of 25\,\unit{\%}
      regarding ice rheological parameters and 14\,\unit{\%} regarding basal
      sliding parameters (Fig.~\ref{fig:sens_ts};
      Table~\ref{tab:sens_extrema}).

@@

      During MIS~4, the modelled sea level potential peaks at
      7.43\,\unit{m}\,s.l.e. (57.6\,\unit{ka}) in the GRIP simulation and at
      4.84\,\unit{m}\,s.l.e. (61.9\,\unit{ka}) in the EPICA simulation,
      corresponding to respectively 86 and 57\,\unit{\%} of modelled MIS~2
      ice volumes.
      These estimates appear little sensitive to the range of parameter
      values tested in our sensitivity study (Sect.~\ref{sec:sens}), which
      yielded relative errors of 12\,\unit{\%} to ice rheological parameters
      and 19\,\unit{\%} to basal sliding parameters
      (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

      The modelled sea level potential corresponding to these persistent
      ice-dispersal centres attains a~minimum of 1.54\,\unit{m}\,s.l.e.
      (42.9\,\unit{ka}) in the GRIP simulation and of 2.55\,\unit{m}\,s.l.e.
      (52.4\,\unit{ka}) in the EPICA simulation, corresponding to
      respectively 18 and 30\,\unit{\%} of the MIS~2 ice volumes.
      However, these numbers should be considered with caution as our
      sensitivity study (Sect.~\ref{sec:sens}) shows that minimum ice volume
      during MIS~3 is highly sensitive to ice flow parameters, with relative
      errors of 45\,\unit{\%} to the range of ice rheological parameters
      tested and 109\,\unit{\%} to the range of basal sliding parameters
      tested (Fig.~\ref{fig:sens_ts}; Table~\ref{tab:sens_extrema}).

@@

\begin{table*}
\caption{%
      Parameter values used in the sensitivity test.}
\label{tab:sens_params}
%\centering\makebox[\textwidth]
{\begin{tabular}{l|ccccc|cc|cc}
    \tophline
            & \multicolumn{5}{c|}{Rheology}
            & \multicolumn{2}{c|}{Sliding}
            & \multicolumn{2}{c}{GRIP scaling} \\
    Config. & $A_{\mathrm{c}}$ & $A_{\mathrm{w}}$
            & $Q_{\mathrm{c}}$ & $Q_{\mathrm{w}}$
            & $E_{\text{SIA}}$ & $\delta$ & $W_{\text{max}}$
            & $f$ & $T_{[32, 22]}$ \\
            & \multicolumn{2}{c}{(\unit{Pa^{-3}\,s^{-1}})}
            & \multicolumn{2}{c}{(\unit{J\,mol^{-1}})}
            & & & (\unit{m}) \\
    \middlehline
    Default$^1$  & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.02 & 2 & 0.38 & 6.2 \\
    \cline{1-10}
    Soft ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 5 & 0.02 & 2 & 0.41 & 6.65 \\
    Hard ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 1 & 0.02 & 2 & 0.36 & 5.95 \\
    \cline{1-10}
    Soft bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.01 & 1 & 0.40 & 6.55 \\
    Hard bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.05 & 5 & 0.36 & 5.85 \\
    \bottomhline
\end{tabular}}
\belowtable{%
      After $^1$\citet{Paterson.Budd.1982,Bueler.Pelt.2015};
      and $^2$\citet{Cuffey.Paterson.2010}.}
\end{table*}

@@

\begin{table*}[p]
\caption{%
      Extremes in Cordilleran ice sheet grounded ice extent and sea-level
      relevant ice volume corresponding to
      MIS~4, 3 and 2 using the GRIP paleo-climate forcing with each
      parameter configuration (Fig.~3). Relative differences (R. diff.) give
      rough error estimates related to varying selected ice rheology and
      basal sliding parameters (Table~\ref{tab:sens_params}).}
\label{tab:sens_extrema}
%\centering\makebox[\textwidth]
{\begin{tabular}{l*{3}{|ccc}}
    \tophline
             & \multicolumn{3}{c|}{Age (ka)}
             & \multicolumn{3}{c|}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m\,s.l.e.)} \\
    Config.  &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    Default  &  57.59 &  42.91 &  19.14
             &   1.93 &   0.67 &   2.09
             &   7.43 &   1.54 &   8.62 \\
    \cline{1-10}
    Soft ice &  58.89 &  49.95 &  21.57
             &   1.99 &   0.54 &   2.11
             &   6.69 &   1.04 &   7.09 \\
    Hard ice &  57.31 &  42.91 &  19.14
             &   1.85 &   0.71 &   2.09
             &   7.56 &   1.74 &   9.22 \\
    R. diff. &    3\,\unit{\%} &   16\,\unit{\%} &   13\,\unit{\%}
             &    7\,\unit{\%} &   25\,\unit{\%} &    1\,\unit{\%}
             &   12\,\unit{\%} &   45\,\unit{\%} &   25\,\unit{\%} \\
    \cline{1-10}
    Soft bed &  58.90 &  49.70 &  21.57
             &   1.88 &   0.55 &   2.12
             &   6.42 &   1.08 &   7.89 \\
    Hard bed &  57.32 &  42.95 &  19.14
             &   1.90 &   0.93 &   2.10
             &   7.84 &   2.77 &   9.10 \\
    R. diff. &    3\,\unit{\%} &   16\,\unit{\%} &   13\,\unit{\%}
             &    1\,\unit{\%} &   57\,\unit{\%} &    1\,\unit{\%}
             &   19\,\unit{\%} &  109\,\unit{\%} &   14\,\unit{\%} \\
    \bottomhline
\end{tabular}}
\end{table*}
